---
title: "GSEAdv: Simulate"
abstract: >
  Shows the methods and functions used to create new gene sets.
date: "`r BiocStyle::doc_date()`"
package: "`r BiocStyle::pkg_ver('GSEAdv')`"
output:
  BiocStyle::html_document:
    fig_caption: true
    toc_float:
      collapsed: true
      toc_depth: 3
author:
- name: Lluís Revilla
  affiliation: 
    - August Pi i Sunyer Biomedical Research Institute (IDIBAPS); Liver Unit, Hospital Clinic
  email: lrevilla@clinic.cat
vignette: >
  %\VignetteIndexEntry{GSEAdv: Describe}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}  
---
```{r knitsetup, message=FALSE, warning=FALSE, include=FALSE}
knitr::opts_knit$set(root.dir = ".")
knitr::opts_chunk$set(collapse = TRUE, warning = TRUE)
BiocStyle::markdown()
library("BiocStyle")
library("GSEAdv")
fl <- system.file("extdata", "Broad.xml", package = "GSEABase")
gss <- getBroadSets(fl)
```

# Introduction

This vignette assumes that the reader is familiar with the infraestructure build in Bioconductor around gene sets, mainly `r Biocpkg("GSEABase")`. It provides examples of how to use the package to create and modify different gene sets.

# Modifying existing gene sets

The first step towards simulate a `GeneSetCollection` is to modify it. There are two ways do modifying them: adding new relationships, or removing genes and `GeneSets`.

## Modifying relationships
### Adding new ones

We might want to add a gene to a gene set or to add a new gene set. We can do that with:

```{r}
add(gss, gene = c("ha", "ZNF474", "CCDC100"), pathway = "new")
add(gss, gene = "ZNF474", pathway = c("chr16q24", "chr5q23"))
```
### Removing

## Removing genes and `GeneSets`
Or we might want to remove a pathway or some genes:

```{r}
# Remove a pathway
drop(gss, pathway = "chr5q23")
drop(gss, pathway = 1) # 1 random pathway
drop(gss, pathway = c("chr5q23", "chr16q24")) 
# Remove some genes
drop(gss, gene = c("AFG3L1", "ANKRD11")) 
drop(gss, gene = 5) # 5 random genes
# Remove both, genes and pathways
drop(gss, pathway = "chr5q23", gene = c("AFG3L1", "ANKRD11"))
drop(Info, pathway = "1430728", gene = c("10", "9"))
```
Dropping by number is aimed to make it possible to bootstrap with different `GeneSetCollections`  comming from the same original one. 

# Simulating 

All the functions to simulate new `GeneSetCollection`s start with `from*`. As a summary of the methods available and what kind of data they use is provided below:

method   | nGenes | nPathways | genesPerPathway| pathwaysPerGene | sizeGenes | sizePathways
------- | ----- | ----- | ----- | ----- | ----- | ----- 
fromGPP |  | (x) | x |  |   |   
fromPPG | (x) |  |  |   x    |   |  
fromGPP_nGenes | x | (x) | x    |  |  |  
fromPPG_nPathways | (x) | x | x    |  |  |  
fromPPG_GPP | (x) | (x) | x | x |     |     
fromSizeGenes | (x) | (x) | (x) | (x) | x    |  
fromSizePathways | (x) | (x) | (x) | (x) |     |  x


: (\#tab:table) The x indicates that it is a required argument of the functions. The parenthesis indicate that it is implicit.

When using more information to the simulations it takes more time (as it is randomly assigned). Also note that they must obey some simple rules: 
 - a gene set cannot be of only one gene
 - a gene cannot be duplicated in a single gene set
 - there can't be duplicated gene sets (gene sets with the same genes)
 
 
```{r fromGPP, eval=FALSE, include=FALSE}
fromGPP(genesPerPathway(Info))
fromPPG(pathwaysPerGene(Info))
```

```{r from_n, eval=FALSE, include=FALSE}
fromPPG_nPathways(pathwaysPerGene(Info), nPathways(Info))
fromGPP_nGenes(genesPerPathway(Info), nGenes(Info))
```

# Side notes

Footnotes are displayed as side notes on the right margin^[this is a side note entered as a footnote], which has the advantage that they appear close to the place where they are defined.


# Session info {.unnumbered}

```{r sessionInfo, echo=FALSE}
sessionInfo()
```
