---
title: "GSEAdv: Evaluate"
abstract: >
  Shows the methods and functions used to evaluate the relationship between genes and gene sets.
date: "`r BiocStyle::doc_date()`"
package: "`r BiocStyle::pkg_ver('GSEAdv')`"
output:
  BiocStyle::html_document:
    fig_caption: true
    toc_float:
      collapsed: true
      toc_depth: 3
author:
- name: Lluís Revilla
  affiliation: 
    - August Pi i Sunyer Biomedical Research Institute (IDIBAPS); Liver Unit, Hospital Clinic
  email: lrevilla@clinic.cat
vignette: >
  %\VignetteIndexEntry{GSEAdv: Describe}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}  
---
```{r knitsetup, message=FALSE, warning=FALSE, include=FALSE}
knitr::opts_knit$set(root.dir = ".")
knitr::opts_chunk$set(collapse = TRUE, warning = TRUE)
BiocStyle::markdown()
library("BiocStyle")
library("GSEAdv")
```
# Introduction

This vignette assumes that the reader is familiar with the infraestructure build in Bioconductor around gene sets, mainly `r Biocpkg("GSEABase")`. It provides examples of how to use the package to understand the relationships between genes and gene sets.

# Distribution of genes and pathways by size:

This functions calculate for each gene or pathway how are their elements distributed:

```{r}
sizeGenes(Info)
sizePathways(Info)
```
We can see that genes are distributed in pathways of size 2 or 3, but pathways have genes in either 1, 2 or 4 pathways. If we are only interested in how many different sizes they are distributed we can use the following functions:

# Number of different sizes per genes or pathways

These two functions evaluate how many elements of different length they are involved in

```{r}
sizesPerGene(Info)
sizesPerPathway(Info)
```

# Style macros

_BiocStyle_ introduces the following macros for referring to _R_ packages:

* `r Biocpkg("IRanges")`, for _Bioconductor_ software, annotation and experiment data packages,
* `r CRANpkg("data.table")`, for _R_ packages available on CRAN,
* `r Githubpkg("rstudio/rmarkdown")`, for _R_ packages available on GitHub,
* `r Rpackage("MyPkg")`, for _R_ packages that are _not_ available on _Bioconductor_, CRAN or GitHub.


# Figures

Assign captions to figures in the code chunk option `fig.cap` to automatically number them, and to be able to reference them, see Figure \@ref(fig:plot). The figure label is generated from the code chunk label by prefixing it with `fig:`.

```{r plot, fig.cap="Regular figure. The first sentence of the figure caption is automatically emphasized to serve as figure title.", echo=FALSE}
plot(cars)
```

Small and wide figures can be specified by `fig.small` and `fig.wide` code chunk options.

```{r small, fig.cap="Small figure. A plot produced by a code chunk with option `fig.small = TRUE`.", fig.small=TRUE, echo=FALSE}
plot(cars)
```

```{r wide, fig.cap="Wide figure. A plot produced by a code chunk with option `fig.wide = TRUE`.", fig.wide=TRUE, echo=FALSE}
plot(cars)
```


# Equations

To number and reference equations, put them in equation environments and assign labels to them, see Equation \@ref(eq:binom).

\begin{equation}
  f\left(k\right) = \binom{n}{k} p^k\left(1-p\right)^{n-k}
  (\#eq:binom)
\end{equation}


# Tables

Like figures, tables with captions will also be numbered and can be referenced, see Table \@ref(tab:table).

Fruit   | Price
------- | -----
bananas | 1.2
apples  | 1.0
oranges | 2.5

: (\#tab:table) A simple table. With caption.


# Cross-references

Apart from referencing figures (Section \@ref(figures)), tables (Section \@ref(tables)), and equations (Section \@ref(equations)), you can also use the same syntax to refer to sections by their default labels generated by pandoc.


# Side notes

Footnotes are displayed as side notes on the right margin^[this is a side note entered as a footnote], which has the advantage that they appear close to the place where they are defined.


# Session info {.unnumbered}

```{r sessionInfo, echo=FALSE}
sessionInfo()
```
