---
title: "About GSEAdv"
abstract: >
  Describes the background of the package, important functions defined in the
  package and some of the applications and usages.
date: "`r BiocStyle::doc_date()`"
package: "`r BiocStyle::pkg_ver('GSEAdv')`"
output:
  BiocStyle::html_document:
    fig_caption: true
    toc_float:
      collapsed: true
      toc_depth: 3
author:
- name: Lluís Revilla
  affiliation: 
    - August Pi i Sunyer Biomedical Research Institute (IDIBAPS); Liver Unit, Hospital Clinic
  email: lrevilla@clinic.cat
vignette: >
  %\VignetteIndexEntry{About GSEAdv}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}  
---
```{r knitsetup, message=FALSE, warning=FALSE, include=FALSE}
knitr::opts_knit$set(root.dir = ".")
knitr::opts_chunk$set(collapse = TRUE, warning = TRUE)
BiocStyle::markdown()
library("BiocStyle")
library("GSEAdv")
```

# Introduction

Bioconductor provides with the package `r Biocpkg("GSEABase")` which provides the basic methods and classes to join, create, and manipulate the gene sets. However few (if any) packages take into account the relationship between genes and gene sets when using them. This package builds upon `GeneSetCollection` providing methods to *describe* the relationship between genes and gene sets, *evaluate* the relationship between genes and gene sets, and *simulate* gene set collections.

# Installation

You can install it from [github](https://www.github.com/llrs/GSEAdv) to get the latest version:
```{r, eval = FALSE}
devtools::install_github("llrs/GSEAdv")
```

or from Bioconductor for a stable release:
```{r Bioc-installation, eval = FALSE}
## try http:// if https:// URLs are not supported
## source("https://bioconductor.org/biocLite.R")
## biocLite("GSEAdv")
```

# Semantics

Thorough the vignettes and the help page a group of genes is both refered as gene sets or as pathways. A pathway or a gene set is defined as a group of genes (more than one) which can't contain twice the same gene. 

Also all the genes should use the same identifier type, Entrez, or a Symbol or Ensembl.

In order of doing this there is a function `check` that checks if the `GeneSetCollection` adhers to this principles:

```{r check}
fl <- system.file("extdata", "Broad.xml", package = "GSEABase")
gss <- getBroadSets(fl)
check(gss)
```


# Usage

Once installed you just need to load it[^1]:

[^1]:To avoid a long message due to GSEAdv's dependencies it is recommend to use `suppressPackageStartupMessages` when loading this package.

```{r load}
library("GSEAdv")
```


The package is thought for three main cases, each one has its own vignette:

 - *Description*
 To provide insight about the relationship between genes and gene sets.
 - *Evaluation*
 Compare and estimate the relationship between genes and gene sets.
 - *Simulate*
 Given some approximations of the relationships between genes and gene sets creates a new gene set collection.


