---
title: "GSEAdv: Evaluate"
abstract: >
  Shows the methods and functions used to evaluate the relationship between genes and gene sets.
date: "`r BiocStyle::doc_date()`"
package: "`r BiocStyle::pkg_ver('GSEAdv')`"
output:
  BiocStyle::html_document:
    fig_caption: true
    toc_float:
      collapsed: true
      toc_depth: 3
author:
- name: Lluís Revilla
  affiliation: 
    - August Pi i Sunyer Biomedical Research Institute (IDIBAPS); Liver Unit, Hospital Clinic
  email: lrevilla@clinic.cat
vignette: >
  %\VignetteIndexEntry{GSEAdv: Describe}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}  
---
```{r knitsetup, message=FALSE, warning=FALSE, include=FALSE}
knitr::opts_knit$set(root.dir = ".")
knitr::opts_chunk$set(collapse = TRUE, warning = TRUE)
BiocStyle::markdown()
library("BiocStyle")
library("GSEAdv")
fl <- system.file("extdata", "Broad.xml", package = "GSEABase")
gss <- getBroadSets(fl)
```

# Introduction

This vignette assumes that the reader is familiar with the infraestructure build in Bioconductor around gene sets, mainly `r Biocpkg("GSEABase")`. 
It provides examples of how to use the package to understand the relationships between genes and gene sets. 
The reader should be also familiarized with the vignette `r Biocpkg("GSEAdv", "vignette_describe.html", "describing")` the relationship between genes and gene sets.

# Isolation

If we want to check if a GeneSetCollection is not a group of isolated genes without relation between different gene sets we can use the function `isolation`

```{r isolation}
isolation(gss)
isolation(Info)
```


# Distribution of genes and pathways by size:

This functions calculate for each gene or pathway how are their elements distributed:

```{r sizes}
sizeGenes(Info)
sizePathways(Info)
```
We can see that genes are distributed in pathways of size 2 or 3, but pathways have genes in either 1, 2 or 4 pathways. If we are only interested in how many different sizes they are distributed we can use the following functions:

# Diferences in genes or pathways 

These two functions evaluate how many elements of different length they are involved in

```{r sizesPer}
sizesPerGene(Info)
sizesPerPathway(Info)
sizesPerPathway(gss)
```

`sizesPerGene` is the equivalent of counting how many rows are not empty on the `sizeGenes` matrix, conversely `sizesPerPathway`.

# Relationships between number of genes per pathway and number of pathways per gene

To see what is the conditional probability between the number of pathways gene in n number of pathways are involved $P( pathways_{genes} | genes_{pathway} )$ there is the function `condPerGene`. While the opposite $P( genes_{pathway} | pathways_{genes} )$) can be done with `condPerPathways`
```{r condPer}
condPerGenes(Info)
condPerPathways(Info)
```

# Other vignettes

In case you missed, here I link previous vignettes and a new vignette

 - `r Biocpkg("GSEAdv", "vignette_describe.html", "*Description*")`
 
 To provide insight about the relationship between genes and gene sets.
 
 - `r Biocpkg("GSEAdv", "vignette_simulate.html", "*Simulate*")`
 
 Given some approximations of the relationships between genes and gene sets creates a new gene set collection.

# Session info {.unnumbered}

```{r sessionInfo, echo=FALSE}
sessionInfo()
```
